Best Practices for Developing Sustainable Scientific Software

### 2016-05-19 WSSSPE3/best_practices/main.tex
catching up

\author{Sandra Gesing$^{(1)}$,
Abani Patra$^{(2)}$,
Daniel S.\ Katz$^{(3)}$,
Kyle E.\ Niemeyer$^{(4)}$,
}

%
\thanks{{}$^{(1)}$ Center for Research Computing, University of Notre Dame, Notre Dame, IN, USA; sandra.gesing@nd.edu}
%
\thanks{{}$^{(2)}$ Mechanical and Aerospace Engineering, University at Buffalo, Buffalo, NY, USA; abani@buffalo.edu}
%
\thanks{{}$^{(3)}$ National Center for Supercomputing Applications (NCSA), Graduate School of Library and Information Science, and Electrical and Computer Engineering, University of Illinois, Urbana-Champaign, IL, USA; d.katz@ieee.org}
%
\thanks{{}$^{(4)}$ School of Mechanical, Industrial, and Manufacturing Engineering, Oregon State University, Corvallis, OR, USA; kyle.niemeyer@oregonstate.edu}
%

@@

Sustainability of software can be investigated under many diverse aspects and dimensions: 
economic, technical, environmental and social are only a few. The scope of this paper will focus 
on scientific environments and their implications and the target group are especially developer 
and project manager of scientific software. To set the stage we picked case studies, which have been 
successful to some degree and analyze why they have been successful. In Section~\ref{sec:comm} are 
practices described, which are partly resulting of the means available in and demands resulting of a scientific environment. 
The subsequent section focus on the social aspects of governing and managing sustainable software, 
while~\ref{sec:funding} is concerned with economic aspects. To be able to measure sustainability of 
software, it is important to define metrics as reference and guideline to be able to create such 
software. Additionally, we will present characteristics of available tools (see~\ref{sec:tools}) 
for managing software but also for analysing and delivering metrics of sustainability. Finally, 
we will conclude with an outlook and suggestion to improve sustainability in a scientific context.

>>>>

\author{Sandra Gesing$^{(1)}$,
Abani Patra$^{(2)}$,
Daniel S.\ Katz$^{(3)}$,
}

%
\thanks{{}$^{(1)}$ Center for Research Computing, University of Notre Dame, Notre Dame, IN, USA; sandra.gesing@nd.edu}
%
\thanks{{}$^{(2)}$ Mechanical and Aerospace Engineering, University at Buffalo, Buffalo, NY, USA; abani@buffalo.edu}
%
\thanks{{}$^{(3)}$ National Center for Supercomputing Applications (NCSA), Graduate School of Library and Information Science, and Electrical and Computer Engineering, University of Illinois, Urbana-Champaign, IL, USA; d.katz@ieee.org}
%

@@

Sustainability of software can be investigated under many diverse aspects and dimensions: 
economic, technical, environmental and social are only a few. The scope of this paper will focus 
on scientific environments and their implications and the target group are especially developer 
and project manager of scientific software. To set the stage we picked case studies, which have been 
successful to some degree and analyze why they have been successful. In Section~\ref{sec:comm} are 
practices described, which are partly resulting of the means available in and demands resulting of a scientific environment. 
The subsequent section focus on the social aspects of governing and managing sustainable software, 
while Section~\ref{sec:funding} is concerned with economic aspects. To be able to measure sustainability of 
software, it is important to define metrics as reference and guideline to be able to create such 
software. Additionally, we will present characteristics of available tools (see Section~\ref{sec:tools}) 
for managing software but also for analyzing and delivering metrics of sustainability. Finally, 
we will conclude with an outlook and suggestion to improve sustainability in a scientific context.
